\texttt{Taxonomic supertree construction with}\texttt{\emph{ Incertae sedis}}\texttt{ taxa}

### 2018-04-20 doc/incertae_sedis.tex
Minor edits and compile fixes.

\author{Benjamin D. Redelings and Mark T. Holder} maketitle

@@

The supertree method described by \citet{redelings2017supertree} is
able to use a set of phylogentic estimates and taxonomy to produce a
supertree with clades names according to the taxonomy. However, the
method was not able to accommodate the fact that taxonomies frequently
contain nodes with uncertain placement. These nodes are often labelled
``\emph{incertae sedis}'', which means ``uncertain seat'' in Latin.
The common interpretation of such taxa is that they may not be moved
further towards the root, but may be moved into their sibling taxa.
For example, a genus with a sibling that is a family may be annotated
as ``\emph{incertae familia}'', indicating that it is unclear which
family the genus should be placed in. \emph{Incertae sedis} taxa
frequently occur when a specimen is identified down to a given
taxonomic level, but no further. Extinct taxa are often \emph{incertae
sedis}. OTT is constructed from several source taxonomies, and these
sources include various ways of indicating that the position of a
taxon within the hierarchy is uncertain. OTT uses a series of
``flags'' to annotate these taxa. The supertree of by
\citet{redelings2017supertree} simply pruned these taxa from the
taxonomy and input trees. Thus, the final supertree lacked taxa which
were \emph{incertae sedis}.

Here describe a supertree method so that we can resolve taxonomic
uncertainty by using published phylogenies to place \emph{incertae
sedis} taxa. Developing this method required deciding on a set
operational semantics to be used when interpreting the \emph{incertae
sedis} annotation. These semantics affect the interpretation of the
phylogenetic statements being made by the input taxonomy and the rules
for applying taxonomic names to the final supertree. In addition to
discussing how to interpret in the \emph{incertae sedis} annotation,
we describe the algorithmic changes to the supertree pipeline that
were required to adequately represent the taxonomic uncertainty. The
approach described here allows us to include thousands of new taxa in
our supertree analysis that were previously filtered out. It will also
enable the Open Tree of Life project to include extinct taxa, since
many of these taxa are \emph{incertae sedis}. This makes substantial
progress towards our goal of comprehensive inclusion of known taxa
into the Open Tree of Life summary tree.

@@

Because of the large number (over 2.6 million) of tips in the full
tree, the supertree method of \citet{redelings2017supertree} uses an
approximation approach that relies on decomposition of the full
problem into subproblems. The decomposition partitioning the list of
taxonomy splits $S(\taxonomy_{p})$ into two lists:
$\mathcal{Z}(\taxonomy_{p}),$the list of splits from the taxonomy tree
which are compatible with every rooted split in the phylogenetic
inputs, and $\mathcal{Y}(\taxonomy_{p})$, the list of taxonomic splits
that conflict with at least one phylogenetic input split. Conceptually
the approximate algorithm works by greedily adding all compatible
splits in altered ranked list of splits:
$\mathcal{S}^{\prime}=\mathcal{Z}(\taxonomy_{p})+S(T_{1})+S(T_{2})+\ldots+S(T_{n})+\mathcal{Y}(\taxonomy_{p})$.

In practice, the decomposition allows the compatibility of splits to
be performed on smaller, relabeled trees. This can be done much more
efficiently than if compatibility was checked using the BUILD
algorithm of \citet{AhoSSU1981} on the full leaf set. Thus, the
supertree algorithm first divides the supertree problem into
independent pieces by bisecting trees at the uncontested taxa which
correspond to $Z$, then solves the supertree problem on these pieces,
and finally glues the pieces back together to produce the full tree.

The taxonomy tree, $\taxonomy_{p}$, is derived from OTT. As mentioned
above, OTT not only contains a taxonomic hierarchy, but also contains
taxonomic labels and annotations (referred to as ``flags'') on the
taxa to carry extra information. The entire hiearchy encoded by OTT is
a complete taxonomic tree, $\taxonomy^{\ast}.$ Nodes in
$\taxonomy^{\ast}$ have a one-to-one correspondence with a set of
labels $\mathcal{N}$ that are called taxon names. We will decompose
$\mathcal{N}$ into the taxon names for leaf taxa \emph{$\mathcal{L}$
}and higher taxa $\mathcal{H}$. We will use $\mathcal{U}$ to refer to
the subset ($\mathcal{U}\subseteq\mathcal{N}$) of taxa that are
annotated with the \emph{incertae sedis }property. Other annotations
identify dubious taxa and taxonomic names that are unwanted artifacts
of the process of building OTT. Thus, the taxonomic tree which is used
as an input to the supertree algorithm is pruned version of
$\taxonomy^{\ast}$. For our purposes, the relevant information from
the taxonomy consists of the taxonomy tree $\taxonomy^{\ast}$, the
labels $\mathcal{N}$, and set $\mathcal{U}$ of \emph{incertae sedis
}taxa. In previous work, taxa in $\mathcal{U}$ were all pruned from
the taxonomic hierarchy when $\taxonomy_{p}$ is extracted from OTT. In
the present work, we consider supertree approaches that operate on a
taxonomic tree $\taxonomy,$which is produced from
$\taxonomy^{\ast}$but does not prune \emph{incertae sedis} taxa.

@@

In order to incorporate \emph{incertae sedis }taxa into a supertree
analyses, we must arrive at an operational definition of the meaning
of the \emph{incertae sedis} label. The core meaning of the phrase is
that the annotated taxon may actually be a member of one of the taxa
that appear as siblings to it in the heirarchical representation of
the taxonomy. In the most expansive interpretation, the author of the
taxonomy is stating that the \emph{incertae sedis} taxon could be
placed anywhere inside one of the sibling taxa or their descendants.
However, frequently only a subset of the possible placement points
would be viewed as plausible. Ideally, the \emph{incertae sedis}
annotation some taxon \emph{x} would be accompanied by a specific
statement about which other taxa might include \emph{x}. Detailed
readings of the character argumentation used by taxonomists could
provide this sort of specific information in many cases, but we lack
such information on the scale of the entire OTT. Here we consider a
few operational approaches for handling the \emph{incertae sedis}
label.

@@

As mentioned above, the incertae sedis may mean that a taxon is
unplaced, but only a few placements within the taxonomy are possible.
For example, a genus may not be characterized as incertae sedis at the
level of the subfamily. Thus, it could be placed inside any of the
hypothesized sub-families, or within its own subfamily, but it would
be unexpected for it to be placed inside one of the other genera
within a family.

In principle this information is easy to incorporate. If
$\mathcal{P}$is the set of taxa that are potential ancestors for
incertae sedis taxon $x$, then $\mathcal{L}(x)$ will occur in neither
the include nor the exclude set of any taxon in $\mathcal{P}$.
Unfortunately, utilizing taxon-specific interpretations of the
incertae sedis flag relies on character argumentation (which is
lacking in OTT) or rank-based information (which is frequently
untrustworthy in OTT). Thus, here we pursue an approach of ascribing a
meaning to the \emph{incertae sedis} label which attempts to capture
the core of the idea that it articulates, but which can be applied
automatically across the taxonomy without additional information.

@@

However, in the \emph{incertae sedis} framework, it is possible for
one name to apply to multiple nodes. For example, in Figure \ref{fig
:One-name-can}b, the name $A$ can apply to nodes $x$ and $y$. Here the
name $A$ corresponds to the split $A1\,A2\,|\bullet B\,C$, leaving out
$D$ since it is \emph{incertae sedis}. The two nodes $x$ and $y$
display the splits $A1\,A2\,|\bullet B\,C\,D$ and $A1\,A2\,D|\bullet
B\,C$ respectively, and both of these splits display the split
$A1\,A2\,|\bullet B\,C$, so the name $A$ can apply to both $x$ and
$y$. This cannot happen without\emph{ incertae sedis} taxa except at
monotypic nodes. When faced with a choice about where to place a name,
our solution is to find the most tip-ward node where the name can
apply and attach the name to this node. This corresponds to a
conservative choice about whether \emph{incertae sedis} taxa should be
added to an existing taxon.
\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/name1nodes2/tax-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/name1nodes2/solution-shaded}}
\hfill{}
\caption{\label{fig:One-name-can}One name can be consistent with multiple nodes.}
\end{figure}

It is also possible for multiple names to apply to a single node. For
example, in Figure \ref{fig:Multiple-names-can}a, the taxon $A$
corresponds to the split $B1\,B2\,C\,|\bullet\,Y$, and its child taxon
$B$ corresponds to the split $B1\,B2\,|\bullet\,Y$. The edge leading
to $A$ is consistent with the split for $B$, but the name $B$ is
applied to the node with the smallest include group. However, in the
solution tree (Fig. \ref{fig:Multiple-names-can}b), there is is only
node for both names $A$ and $B$ to apply to. In this case, we solve
this problem by introducing a monotypic parent, and applying $A$ to
the newly created parent node (Fig. \ref{fig:Multiple-names-can}c).
\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/tax-shaded}}
\hfill{}subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/synth-shaded}}
\hfill{}subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/solution2-shaded}}
\hfill{}
\caption{\label{fig:Multiple-names-can}Multiple names can apply to a single node.}
\end{figure}

@@

After we have attached taxon names to the synthesis tree, we would
like to interpret the position of these names in terms of placing
\emph{incertae sedis }taxa in a revised taxonomy. This would allow us
to interpret the synthesis tree as saying that phylogenetic
information has (for example) placed genus A within family B, or
perhaps outside of all named families. The simplest approach to
placement involves noting whenever a taxon $B$ is a descendant of a
taxon $A$ on the named synthesis tree but not the taxonomy tree. For
example, in Figure \ref{fig:Handling-incertae-sedis}, taxon $C$ is a
descendant of taxon $A$ in the synthetic tree (Fig. \ref{fig:Handling-
incertae-sedis}d), but not the taxonomy (Fig. \ref{fig:Handling-
incertae-sedis}b). Thus, we could say that the synthetic tree places
$C$ within $A$. Furthermore, the synthetic tree additionally places
$C$ within $AB$.

A slightly more general case is shown in Figure \ref{fig:Broken-
incertae-sedis}. Here taxon $C1$ is an descendant of $A$ in the
synthesis tree, but not the taxonomy. Thus $C1$ is placed within $A$.
We note that the most recent common ancestor (MRCA) of $A$ and $C1$ in
the taxonomy is the root node, and the path to the MRCA from $A$ is
$A\to AB\to root$, while the path to the MRCA from $C1$ is $C1\to C\to
root$. Here the additional node $AB$ on the path to the MRCA from $A$
indicates that $C1$ is placed within $AB$ as well as $A$. In contrast,
the additional node $C$ on the path to the MRCA indicates that $C1$ is
a broken \emph{incertae sedis} taxon. Thus we may consider an
algorithm that, for each taxon $B$ on the synthesis tree, finds the
closest ancestral taxon $A$, and checks if $A$ is an ancestor of $B$
on the taxonomy. If not, we compute the paths $A\to A_{1}\to\ldots\to
MRCA(A,B)$ and $B\to B_{1}\to\ldots\to MRCA(A,B)$.

However, this assumes that \emph{incertae sedis} taxa are always
placed within their siblings, or descendants of their siblings. As
such, it does not handle cases such as Figure \ref{fig:Placing-one-
incertae}, in which $I$ is first placed as sibling to $AI$, and then
$AI$ is placed within $I$. Note that $I$ is not a sibling of $AI$ or a
descendant of a sibling. Furthermore, the MRCA of $I$ and $AI$ is the
root, and the path from $AI$ to the root includes $A$, but $A$ is not
broken. We suggest that this conundrum may be resolved by first
processing the placement of $I$ within $A$ by modifying the taxonomy
to reflect this placement\footnote{Does this actually work? Some
placements break taxa.}. When we consider the placement of $AI$ within
$I$ on this modified taxonomy, the MRCA of $I$ and $AI$ is now $A$
instead of the root, and the correct interpretation results. We
suggest that by walking the tree in a pre-order fashion and processing
shallowing placements first, the previous approach of finding the MRCA
and labelling intermediate nodes as either additional placements or
broken incertae sedis (or incertae\_sedis\_inherited) taxa holds. Such
an approach should also handle cases containing nested incertae sedis
taxa, as in Figure \ref{fig:Placing-one-incertae}.

@@

We therefore seek a synthesis method that can correctly introduce
\emph{incertae sedis} taxa into containing taxa without breaking the
containing taxa. We change the semantics of names to imply, not the
exclusion of all non-included taxa, but only some non-included taxa,
as described in section \ref{sec:Semantics-of- incertae}. As a result
of this change in semantics, placing an IS taxon $A$ within a sister
taxon $B$ no longer results in conflict with $B$. This allows us to
retain the split for $B$ within the synthesis tree, so that taxonomy-
only children of $B$ are correctly grouped with their siblings that
are referenced by the input trees. We may then retain the name for the
no-longer-broken taxon. Finally, we are then able to stop filtering
\emph{incertae sedis} taxa, so that they appear in the synthesis tree.
Thus, the synthesis tree is able to represent substantially more
species, without suffering the loss of taxa and the loss of structure.

@@

{[}\textbf{BDR:} \emph{maybe we should merge this with Figure.
\ref{fig:Broken- incertae-sedis}. Although the point of figure
\ref{fig:placement-example1} is to talk about sub-problem
decomposition as well as what the result looks like.}{]}

The addition of \emph{incertae sedis} taxa allows new types of
conflict between input trees. For example, different input trees might
place an incertae sedis taxon in conflicting locations. This is
illustrated in Figure \ref{fig :placement-example1}, where the IS
taxon (ott7,ott6)ott10 is placed as sister to ott1 by phylogeny
$\tau_{1}$ and as sister ott3 by phylogeny $\tau_{2}$.

@@

\subsection{Case 1: An emph{incertae sedis} clade}

When a clade is marked as \emph{incertae sedis}, we need to consider
two cases. In the first case, the clade may be placed within a sister
clade intact (Figure \ref{fig:Handling-incertae-sedis}). However, if
the clade is not monophyletic in the synthesis tree, then we allow the
members of the clade to be placed separately (Figure \ref{fig:Broken-
incertae-sedis}).

>>>>

\author{Benjamin D. Redelings and Mark T. Holder}

\maketitle

@@

The supertree method described by \citet{redelings2017supertree} is
able to use a set of phylogenetic estimates and taxonomy to produce a
supertree with clades names according to the taxonomy. However, the
method was not able to accommodate the fact that taxonomies frequently
contain nodes with uncertain placement. These nodes are often labelled
``\emph{incertae sedis}'', which means ``uncertain seat'' in Latin.
The common interpretation of such taxa is that they may not be moved
further towards the root, but may be moved into their sibling taxa.
For example, a genus with a sibling that is a family may be annotated
as ``\emph{incertae familia}'', indicating that it is unclear which
family the genus should be placed in. \emph{Incertae sedis} taxa
frequently occur when a specimen is identified down to a given
taxonomic level, but no further. Extinct taxa are often \emph{incertae
sedis}. OTT is constructed from several source taxonomies, and these
sources include various ways of indicating that the position of a
taxon within the hierarchy is uncertain. OTT uses a series of
``flags'' to annotate these taxa. The supertree of by
\citet{redelings2017supertree} simply pruned these taxa from the
taxonomy and input trees. Thus, the final supertree lacked taxa which
were \emph{incertae sedis} or the descendants of such taxa.

Here describe a supertree method so that we can resolve taxonomic
uncertainty by using published phylogenies to place \emph{incertae
sedis} taxa. Developing this method required deciding on a set
operational semantics to be used when interpreting the \emph{incertae
sedis} annotation. These semantics affect the interpretation of the
phylogenetic statements being made by the input taxonomy and the rules
for applying taxonomic names to the final supertree. In addition to
discussing how to interpret the \emph{incertae sedis} annotation,
we describe the algorithmic changes to the supertree pipeline that
were required to adequately represent the taxonomic uncertainty. The
approach described here allows us to include thousands of new taxa in
our supertree analysis that were previously filtered out. It will also
enable the Open Tree of Life project to include extinct taxa, since
many of these taxa are \emph{incertae sedis}. This makes substantial
progress towards our goal of comprehensive inclusion of known taxa
into the Open Tree of Life summary tree.

@@

Because of the large number (over 2.6 million) of tips in the full
tree, the supertree method of \citet{redelings2017supertree} uses an
approximation approach that relies on decomposition of the full
problem into subproblems. The decomposition partitions the list of
taxonomy splits $S(\taxonomy_{p})$ into two lists:
$\mathcal{Z}(\taxonomy_{p}),$ the list of splits from the taxonomy tree
which are compatible with every rooted split in the phylogenetic
inputs, and $\mathcal{Y}(\taxonomy_{p})$, the list of taxonomic splits
that conflict with at least one phylogenetic input split. Conceptually
the approximate algorithm works by greedily adding all compatible
splits in altered ranked list of splits:
$\mathcal{S}^{\prime}=\mathcal{Z}(\taxonomy_{p})+S(T_{1})+S(T_{2})+\ldots+S(T_{n})+\mathcal{Y}(\taxonomy_{p})$.

In practice, the decomposition allows the compatibility of splits to
be checked on smaller, relabeled trees. This can be done much more
efficiently than if compatibility was checked using the BUILD
algorithm of \citet{AhoSSU1981} on the full leaf set. Thus, the
supertree algorithm first divides the supertree problem into
independent pieces by bisecting trees at the uncontested taxa which
correspond to $Z$, then solves the supertree problem on these pieces,
and finally glues the pieces back together to produce the full tree.

The taxonomy tree, $\taxonomy_{p}$, is derived from OTT. As mentioned
above, OTT not only contains a taxonomic hierarchy, but also contains
taxonomic labels and annotations (referred to as ``flags'') on the
taxa to carry extra information. The entire hiearchy encoded by OTT is
a complete taxonomic tree, $\taxonomy^{\ast}.$ Nodes in
$\taxonomy^{\ast}$ have a one-to-one correspondence with a set of
labels $\mathcal{N}$ that are called taxon names. We will decompose
$\mathcal{N}$ into the taxon names for leaf taxa \emph{$\mathcal{L}$
}and higher taxa $\mathcal{H}$. We will use $\mathcal{U}$ to refer to
the subset ($\mathcal{U}\subseteq\mathcal{N}$) of taxa that are
annotated with the \emph{incertae sedis }property. Other annotations
identify dubious taxa and taxonomic names that are unwanted artifacts
of the process of building OTT. Thus, the taxonomic tree which is used
as an input to the supertree algorithm is a pruned version of
$\taxonomy^{\ast}$. For our purposes, the relevant information from
the taxonomy consists of the taxonomy tree $\taxonomy^{\ast}$, the
labels $\mathcal{N}$, and set $\mathcal{U}$ of \emph{incertae sedis
}taxa. In previous work, taxa in $\mathcal{U}$ and all of their descendants were all pruned from
the taxonomic hierarchy when $\taxonomy_{p}$ is extracted from OTT. In
the present work, we consider supertree approaches that operate on a
taxonomic tree $\taxonomy$, which is produced from
$\taxonomy^{\ast}$but does not prune \emph{incertae sedis} taxa.

@@

In order to incorporate \emph{incertae sedis }taxa into a supertree
analyses, we must arrive at an operational definition of the meaning
of the \emph{incertae sedis} label. The core meaning of the phrase is
that the annotated taxon may actually be a member of one of the taxa
that appear as siblings to it in the heirarchical representation of
the taxonomy. In the most expansive interpretation, the author of the
taxonomy is stating that the \emph{incertae sedis} taxon could be
placed anywhere inside one of the sibling taxa or their descendants.
However, frequently only a subset of the possible placement points
would be viewed as plausible. Ideally, the \emph{incertae sedis}
annotation for some taxon \emph{x} would be accompanied by a specific
statement about which other taxa might include \emph{x}. Detailed
readings of the character argumentation used by taxonomists could
provide this sort of specific information in many cases, but we lack
such information on the scale of the entire OTT. Here we consider a
few operational approaches for handling the \emph{incertae sedis}
label.

@@

As mentioned above, the \emph{incertae sedis} annotation may mean that a taxon is
unplaced, but only a few placements within the taxonomy are possible.
For example, a genus may not be characterized as incertae sedis at the
level of the subfamily. Thus, it could be placed inside any of the
hypothesized sub-families, or within its own subfamily, but it would
be unexpected for it to be placed inside one of the other genera
within a family.

In principle this information is easy to incorporate. If
$\mathcal{P}$ is the set of taxa that are potential ancestors for
incertae sedis taxon $x$, then $\mathcal{L}(x)$ will occur in neither
the include nor the exclude set of any taxon in $\mathcal{P}$.
Unfortunately, utilizing taxon-specific interpretations of the
incertae sedis flag relies on character argumentation (which is
lacking in OTT) or rank-based information (which is frequently
untrustworthy in OTT). Thus, here we pursue an approach of ascribing a
meaning to the \emph{incertae sedis} label which attempts to capture
the core of the idea that it articulates, but which can be applied
automatically across the taxonomy without additional information.

@@

However, in the \emph{incertae sedis} framework, it is possible for
one name to apply to multiple nodes. For example, in Figure \ref{fig:One-name-can}b,
the name $A$ can apply to nodes $x$ and $y$. Here the
name $A$ corresponds to the split $A1\,A2\,|\bullet B\,C$, leaving out
$D$ since it is \emph{incertae sedis}. The two nodes $x$ and $y$
display the splits $A1\,A2\,|\bullet B\,C\,D$ and $A1\,A2\,D|\bullet
B\,C$ respectively, and both of these splits display the split
$A1\,A2\,|\bullet B\,C$, so the name $A$ can apply to both $x$ and
$y$. This cannot happen without\emph{ incertae sedis} taxa except at
monotypic nodes. When faced with a choice about where to place a name,
our solution is to find the most tip-ward node where the name can
apply and attach the name to this node. This corresponds to a
conservative choice about whether \emph{incertae sedis} taxa should be
added to an existing taxon.
\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/name1nodes2/tax-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/name1nodes2/solution-shaded}}
\hfill{}
\caption{\label{fig:One-name-can}One name can be consistent with multiple nodes.}
\end{figure}

It is also possible for multiple names to apply to a single node. For
example, in Figure \ref{fig:Multiple-names-can}a, the taxon $A$
corresponds to the split $B1\,B2\,C\,|\bullet\,Y$, and its child taxon
$B$ corresponds to the split $B1\,B2\,|\bullet\,Y$. The edge leading
to $A$ is consistent with the split for $B$, but the name $B$ is
applied to the node with the smallest include group. However, in the
solution tree (Fig. \ref{fig:Multiple-names-can}b), there is is only
node for both names $A$ and $B$ to apply to. In this case, we solve
this problem by introducing a monotypic parent, and applying $A$ to
the newly created parent node (Fig. \ref{fig:Multiple-names-can}c).
\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/tax-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/synth-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/solution2-shaded}}
\hfill{}
\caption{\label{fig:Multiple-names-can}Multiple names can apply to a single node.}
\end{figure}

@@

After we have attached taxon names to the synthesis tree, we would
like to interpret the position of these names in terms of placing
\emph{incertae sedis }taxa in a revised taxonomy. This would allow us
to interpret the synthesis tree as saying that phylogenetic
information has (for example) placed genus A within family B, or
perhaps outside of all named families. The simplest approach to
placement involves noting whenever a taxon $B$ is a descendant of a
taxon $A$ on the named synthesis tree but not the taxonomy tree. For
example, in Figure \ref{fig:Handling-incertae-sedis}, taxon $C$ is a
descendant of taxon $A$ in the synthetic tree (Fig. \ref{fig:Handling-incertae-sedis}d),
but not the taxonomy (Fig. \ref{fig:Handling-incertae-sedis}b).
Thus, we could say that the synthetic tree places
$C$ within $A$. Furthermore, the synthetic tree additionally places
$C$ within $AB$.

A slightly more general case is shown in Figure \ref{fig:Broken-incertae-sedis}.
Here taxon $C1$ is an descendant of $A$ in the
synthesis tree, but not the taxonomy. Thus $C1$ is placed within $A$.
We note that the most recent common ancestor (MRCA) of $A$ and $C1$ in
the taxonomy is the root node, and the path to the MRCA from $A$ is
$A\to AB\to root$, while the path to the MRCA from $C1$ is $C1\to C\to
root$. Here the additional node $AB$ on the path to the MRCA from $A$
indicates that $C1$ is placed within $AB$ as well as $A$. In contrast,
the additional node $C$ on the path to the MRCA indicates that $C1$ is
a broken \emph{incertae sedis} taxon. Thus we may consider an
algorithm that, for each taxon $B$ on the synthesis tree, finds the
closest ancestral taxon $A$, and checks if $A$ is an ancestor of $B$
on the taxonomy. If not, we compute the paths $A\to A_{1}\to\ldots\to
MRCA(A,B)$ and $B\to B_{1}\to\ldots\to MRCA(A,B)$.

However, this assumes that \emph{incertae sedis} taxa are always
placed within their siblings, or descendants of their siblings. As
such, it does not handle cases such as Figure \ref{fig:Placing-one-incertae},
in which $I$ is first placed as sibling to $AI$, and then
$AI$ is placed within $I$. Note that $I$ is not a sibling of $AI$ or a
descendant of a sibling. Furthermore, the MRCA of $I$ and $AI$ is the
root, and the path from $AI$ to the root includes $A$, but $A$ is not
broken. We suggest that this conundrum may be resolved by first
processing the placement of $I$ within $A$ by modifying the taxonomy
to reflect this placement\footnote{Does this actually work? Some
placements break taxa.}. When we consider the placement of $AI$ within
$I$ on this modified taxonomy, the MRCA of $I$ and $AI$ is now $A$
instead of the root, and the correct interpretation results. We
suggest that by walking the tree in a pre-order fashion and processing
shallowing placements first, the previous approach of finding the MRCA
and labelling intermediate nodes as either additional placements or
broken incertae sedis (or incertae\_sedis\_inherited) taxa holds. Such
an approach should also handle cases containing nested incertae sedis
taxa, as in Figure \ref{fig:Placing-one-incertae}.

@@

We therefore seek a synthesis method that can correctly introduce
\emph{incertae sedis} taxa into containing taxa without breaking the
containing taxa. We change the semantics of names to imply, not the
exclusion of all non-included taxa, but only some non-included taxa,
as described in section \ref{sec:Semantics-of-incertae}. As a result
of this change in semantics, placing an IS taxon $A$ within a sister
taxon $B$ no longer results in conflict with $B$. This allows us to
retain the split for $B$ within the synthesis tree, so that taxonomy-
only children of $B$ are correctly grouped with their siblings that
are referenced by the input trees. We may then retain the name for the
no-longer-broken taxon. Finally, we are then able to stop filtering
\emph{incertae sedis} taxa, so that they appear in the synthesis tree.
Thus, the synthesis tree is able to represent substantially more
species, without suffering the loss of taxa and the loss of structure.

@@

{[}\textbf{BDR:} \emph{maybe we should merge this with Figure.
\ref{fig:Broken-incertae-sedis}. Although the point of figure
\ref{fig:placement-example1} is to talk about sub-problem
decomposition as well as what the result looks like.}{]}

The addition of \emph{incertae sedis} taxa allows new types of
conflict between input trees. For example, different input trees might
place an incertae sedis taxon in conflicting locations. This is
illustrated in Figure \ref{fig:placement-example1}, where the IS
taxon (ott7,ott6)ott10 is placed as sister to ott1 by phylogeny
$\tau_{1}$ and as sister ott3 by phylogeny $\tau_{2}$.

@@

\subsection{Case 1: An \emph{incertae sedis} clade}

When a clade is marked as \emph{incertae sedis}, we need to consider
two cases. In the first case, the clade may be placed within a sister
clade intact (Figure \ref{fig:Handling-incertae-sedis}). However, if
the clade is not monophyletic in the synthesis tree, then we allow the
members of the clade to be placed separately (Figure \ref{fig:Broken-incertae-sedis}).
